---
title: "Multivariate Analysis of Network"
author: "Gosuddin Siddiqi"
date: "February 25, 2017"
output: pdf_document
---

```{r setup, include=FALSE, warning= FALSE,message=FALSE}
#load the library
library(statnet)
library(circlize)
load("trade.RData")
```

 This data captures trade in various types of products/materials among countries. Let explore our data.

```{r}
str(trade)

#What kind of trades?
row.names(trade)

#What all countries participate?
colnames(trade)
  

#Visualizing all trade networks
chordDiagram(trade[1,,])  #Minerals

chordDiagram(trade[2,,])  #CRUDE_MATERIALS

chordDiagram(trade[3,,])  #FOODS

chordDiagram(trade[4,,])  #MANUFACTURED_GOODS

chordDiagram(trade[5,,])  #DIPLOMATIC_EXCHANGE

```


##### (a) Clustering #####

Computing a hierarchical clustering of the trade networks, based on the Hamming distance. Comparing this with a two-dimensional MDS solution on the same data.

```{r}
tradehd<-hdist(trade)
tradehd

# Preparing MDS solution
trademds<-cmdscale(tradehd)
trademds
plot(trademds,type="n")
text(trademds,label=rownames(trade))

#plot clusters
tradehc<-hclust(as.dist(tradehd))
plot(tradehc,labels=c("MINERALS","CRUDE_MATERIALS","FOODS","MANUFACTURED_GOODS","DIPLOMATIC_EXCHANGE"))         
rect.hclust(tradehc,k=3)


```



##### (b) PCA #####

Conducting PCA analysis to understand the relationship in the trade network.

```{r}
tradecor <- gcor(trade)
tradecor

#Prepare Eigen
tradeeig <- eigen(tradecor)

evals<-tradeeig$value                                      
evals/sum(evals)  

#Screenplot
barplot(evals/sum(evals),names.arg=1:length(evals))


#load 2 components
load<-tradeeig$vector[,1:2]
rownames(load)<-rownames(trade)
load

#generate plot
plot(load[,1:2],type="n",asp=1,xlab="PC 1",ylab="PC 2")
abline(h=0,v=0,lty=3)
arrows(0,0,load[,1],load[,2],col=2)
text(load[,1:2],label=rownames(trade))

```
The variance in the underlying trade network can be explained sufficiently by the trade of minerals alone. Almost 56% variance is explained by the Minerals alone.

##### (c) Discussion #####

Discussing our PCA results with screen plots.

From the MDS plot we observe that there is something unqiue about Minerals that places it far away from the Diplomatic_Exchange and the cluster of Crude_materials, manufactured good and foods. 
This discussion validates with variation observed in the PCA plot. Minerals suffices for the 56% variance and that Minerals is far away from the rest of the trade networks.
